Supplementary Information: Tropical tree drought sensitivity is jointly shaped by drought characteristics, species adaptations, and individual microenvironments

### 2025-04-10 doc/manuscript_hkk_drought_sensitivity.Rmd
fixed issue with mm calculation

tree.time.sp <- tree.time %>%
  group_by(spfull) %>%
  dplyr::summarise(
    inc_annual_med = round(median(inc_annual, na.rm = T), 2),
    inc_annual_sd = round(sd(inc_annual, na.rm = T), 2)
  ) %>%
  arrange(inc_annual_med)

sens.sp <- tree.time %>%
  filter(yr %in% c(2010, 2015)) %>%
  group_by(spfull, yr) %>%
  dplyr::summarise( # all values in mm
    sens_med = round(median(sens.prop / 10, na.rm = T), 2),
    sens_mean = round(mean(sens.prop / 10, na.rm = T), 2),
    sens_sd = round(sd(sens.prop / 10, na.rm = T), 2)
  ) %>%
  arrange(sens_med)

@@

The increment timeseries using dendrometer band measurements revealed plot- and species-level patterns in interannual growth variation (Fig `r fig_drought_responses`a). 
Median annualised growth rate across the years of measurement for the whole plot was `r round(median(tree.time$inc_annual/10, na.rm=T),2)` $\pm$ `r round(sd(tree.time$inc_annual/10, na.rm=T),2)` mm. 
Median annual increment differed by species with *`r tree.time.sp$spfull[1]`* having the lowest (`r tree.time.sp$inc_annual_med[1]` $\pm$ `r tree.time.sp$inc_annual_sd[1]` mm) and *`r tree.time.sp$spfull[nrow(tree.time.sp)]`* the highest (`r tree.time.sp$inc_annual_med[nrow(tree.time.sp)]` $\pm$ `r tree.time.sp$inc_annual_sd[nrow(tree.time.sp)]` mm) among the `r nrow(tree.time.sp)` species analysed. 
*(See [issue #27](https://github.com/forestgeo/growth-precip-thailand/issues/27))*

Drought years were associated with lower growth across the plot. 
In 2010, the median annual increment was `r round(median(tree.time$inc_annual[tree.time$yr==2010]/10, na.rm=T),2)` mm, `r round(-100*((median(tree.time$inc_annual[tree.time$yr==2010]/10, na.rm=T)-median(tree.time$inc_annual/10, na.rm=T))/median(tree.time$inc_annual/10, na.rm=T)),2)`% lower than the median across all years. 
The drought year 2015 had the lowest median annual increment `r round(median(tree.time$inc_annual[tree.time$yr==2015]/10, na.rm=T),2)` mm, `r round(-100*((median(tree.time$inc_annual[tree.time$yr==2015]/10, na.rm=T)-median(tree.time$inc_annual/10, na.rm=T))/median(tree.time$inc_annual/10, na.rm=T)),2)`% lower than the median across all years. 
While median species responses to the 2010 drought were variable, all species had lower increment than a typical year in 2015. 
Besides decrease in magnitude of growth, the proportion of trees with negative growth was also higher in the drought years (Fig S4). 
Across the years, on average, `r round(100*mean(tapply(tree.time$inc_annual, tree.time$yr, function(x) sum(x < 0) / length(x))),2)`% of the trees had negative growth, with a minimum of `r round(100*min(tapply(tree.time$inc_annual, tree.time$yr, function(x) sum(x < 0) / length(x))),2)`% in `r names(sort(tapply(tree.time$inc_annual, tree.time$yr, function(x) sum(x < 0) / length(x))))[1]`. 
The drought year 2015 had the maximum proportion of trees with negative growth in the timeseries, `r round(100*sum(tree.time$inc_annual[tree.time$yr==2015]<0)/nrow(tree.time[tree.time$yr==2015,]),2)`%.
*(See [issue #28](https://github.com/forestgeo/growth-precip-thailand/issues/28))*

>>>>

tree.time.sp <- tree.time %>%
  group_by(spfull) %>%
  dplyr::summarise( # change to mm
    inc_annual_med = round(median(inc_annual * 10, na.rm = T), 2),
    inc_annual_sd = round(sd(inc_annual * 10, na.rm = T), 2)
  ) %>%
  arrange(inc_annual_med)

sens.sp <- tree.time %>%
  filter(yr %in% c(2010, 2015)) %>%
  group_by(spfull, yr) %>%
  dplyr::summarise(
    sens_med = round(median(sens.prop, na.rm = T), 2),
    sens_mean = round(mean(sens.prop, na.rm = T), 2),
    sens_sd = round(sd(sens.prop, na.rm = T), 2)
  ) %>%
  arrange(sens_med)

@@

The increment timeseries using dendrometer band measurements revealed plot- and species-level patterns in interannual growth variation (Fig `r fig_drought_responses`a). 
Median annualised growth rate across the years of measurement for the whole plot was `r round(median(tree.time$inc_annual*10, na.rm=T),2)` $\pm$ `r round(sd(tree.time$inc_annual*10, na.rm=T),2)` mm. 
Median annual increment differed by species with *`r tree.time.sp$spfull[1]`* having the lowest (`r tree.time.sp$inc_annual_med[1]` $\pm$ `r tree.time.sp$inc_annual_sd[1]` mm) and *`r tree.time.sp$spfull[nrow(tree.time.sp)]`* the highest (`r tree.time.sp$inc_annual_med[nrow(tree.time.sp)]` $\pm$ `r tree.time.sp$inc_annual_sd[nrow(tree.time.sp)]` mm) among the `r nrow(tree.time.sp)` species analysed. 
<!-- *(See [issue #27](https://github.com/forestgeo/growth-precip-thailand/issues/27))* -->

Drought years were associated with lower growth across the plot. 
In 2010, the median annual increment was `r round(median(tree.time$inc_annual[tree.time$yr==2010]*10, na.rm=T),2)` mm, `r round(-100*((median(tree.time$inc_annual[tree.time$yr==2010]*10, na.rm=T)-median(tree.time$inc_annual*10, na.rm=T))/median(tree.time$inc_annual*10, na.rm=T)),2)`% lower than the median across all years. 
The drought year 2015 had the lowest median annual increment `r round(median(tree.time$inc_annual[tree.time$yr==2015]*10, na.rm=T),2)` mm, `r round(-100*((median(tree.time$inc_annual[tree.time$yr==2015]*10, na.rm=T)-median(tree.time$inc_annual*10, na.rm=T))/median(tree.time$inc_annual*10, na.rm=T)),2)`% lower than the median across all years. 
While median species responses to the 2010 drought were variable, all species had lower increment than a typical year in 2015. 
Besides decrease in magnitude of growth, the proportion of trees with negative growth was also higher in the drought years (Fig S4). 
Across the years, on average, `r round(100*mean(tapply(tree.time$inc_annual, tree.time$yr, function(x) sum(x < 0) / length(x))),2)`% of the trees had negative growth, with a minimum of `r round(100*min(tapply(tree.time$inc_annual, tree.time$yr, function(x) sum(x < 0) / length(x))),2)`% in `r names(sort(tapply(tree.time$inc_annual, tree.time$yr, function(x) sum(x < 0) / length(x))))[1]`. 
The drought year 2015 had the maximum proportion of trees with negative growth in the timeseries, `r round(100*sum(tree.time$inc_annual[tree.time$yr==2015]<0)/nrow(tree.time[tree.time$yr==2015,]),2)`%.
<!-- *(See [issue #28](https://github.com/forestgeo/growth-precip-thailand/issues/28))* -->
